Issues with Dynamical Measurements Across the Disk

### 2019-02-25 paper/ms.tex
more changes to section 3, and added some aux files to gitignore

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation m12i contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for gas particles around roughly solar positions (with
cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
$98\,\pc$.

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
gas disk is thought to have a scale height of {\bf XXX} pc (XX times the
median gas softening length) and the thin stellar disk a scale height of {\bf
XXX} pc (XX times the stellar softening length). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, and indeed the gas and stellar scale heights of
the simulated galaxies are all somewhat larger than the MW's. However, in this
work we are considering azimuthal variations in the midplane \emph{as defined
by the mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

@@

\subsection{Local Midplane} \label{ssec:local_midplane}
Using the three simulations, we determine the \emph{local} midplane that an
observer might measure if they were situated in each of these galaxies. We
place our imaginary observer at the solar radius of $8.2\,\kpc$ and azimuth
$0<\phi<2\pi$\footnote{Our nearly immortal observer also has warp-drive.}, and
consider stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
perpendicular to the fiducial disk.

This median height is taken to be what our observer would measure as the local
Galactic midplane at each $\phi$, and is reported for each galaxy in
Figure~\ref{fig:midplane}. We find that only $10$ iterations of this procedure
are necessary for convergence. We bootstrap resample $1000$ times and report
the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile as the dashed-line error
bars. This bootstrap resampling is performed by first selecting all stars
within a height of $2\,\kpc$, resampling that selection, and then repeating
the $10$ iterations.

To allow for potential small inaccuracies in the fiducial coordinate systems,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for m12i, m12f, and m12m, respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for m12i, m12f, and m12m. These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
would make as a function of azimuth for each galaxy. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.

@@

\subsection{LSR Variations} \label{ssec:lsr_var}
For similar reasons to the midplane varying across the disk, we also expect
that the local standard of rest (LSR) is a function of azimuth as well. We
perform this calculation in Appendix~\ref{app:lsr}, but in order to perform a
best-fit subtraction as in the previous section is more involved, which we
delay to future work. We do note, however, that the variation in the LSR is
less pronounced than for the midplane, at least for how we calculate the LSR.

>>>>

These simulations contain dark matter particles of mass $\sim35,000\,\Msun$,
gas particles of mass $\sim 7000 \text{--} 20,000\,\Msun$, and star particles
of mass $\sim 5000 \text{--} 7000\,
\Msun$,\footnote{In fact, the simulation m12i contained a gas
splitting bug which causes a small number of gas and star particles to have
larger than typical masses. These constitute only $\sim0.2\%$ of the star
particles in the $\z=0$ snapshots, and are only a factor of a few more massive
than typical star particles. Since our results are broadly consistent between
the three simulations, we ignore this minor complication.} with the lower end
coming from stellar evolution \citep{2018arXiv180610564S}. Softening lengths
for dark matter and star particles are fixed at $112\,\pc$ and $11.2\,\pc$,
respectively.\footnote{This is $2.8$ times the often-quoted
Plummer-equivalent.} The gas softening length is adaptive, but at $z=0$ the
median softening length for gas particles around roughly solar positions (with
cylindrical radii within $500\,\pc$ of $8.2\,\kpc$ and $\abs{z}<1\,\kpc$) is
$98\,\pc$. %***see Issue***

The softening lengths used in the simulations can affect the ability to
resolve the very thinnest planar structures, which in turn can affect how much
the density-based midplane varies as a function of azimuth. The Milky Way's
dense, star-forming gas disk is thought to have a scale height of about 40 pc (XXX times the
median gas softening length; \citealt{2019ApJ...871..145A}) and the thin stellar disk a scale height of about 300 pc (30 times the stellar softening length; \citealt{2008ApJ...673..864J}). We therefore expect that
resolution effects may still be affecting the scale heights of these
components in the simulations, and indeed the gas and stellar scale heights of
the simulated galaxies are equal to or larger than the MW's. However, in this
work we are considering azimuthal variations in the midplane \emph{as defined
by the mass density}, which will subsequently be lower in the simulations than
in the MW: within 200 pc of the Sun the measured gas and stellar volume
densities in the MW are roughly comparable at $\sim .04\ \Msun\
\unit{pc}{-3}$, while in the simulations they are a factor of a few lower (see
Table 3 of \citealt{2018arXiv180610564S}). We thus expect that the degree of
midplane variation we find in the simulated systems is likely to be
\emph{less} than in the real MW, since on the smallest scales the simulations
are \emph{smoother} than the real Galaxy.

@@

\subsection{Local Midplane} \label{ssec:local_midplane}
Using the three simulations, we determine the local midplane that an
observer might measure if they were situated in each of these galaxies, as a function of azimuth at the solar circle. Starting from the coordinate system described in the previous section, which is approximately aligned so that the $z$ coordinate is perpendicular to the disk plane, we
place our imaginary observer at $z=0$ and a solar radius of $8.2\,\kpc$ and vary the azimuth between 
$0<\phi<2\pi$\footnote{Our very long-lived observer also has warp-drive.}. At each value of $\phi$ we then
compute the median $z$ for stars within a cylinder of radius $0.5\,\kpc$ and height $1\,\kpc$
perpendicular to the fiducial disk. We then re-define this median $z$ as the new midplane of the cylinder, re-select stars, and iterate until the median $z$ value converges. We find that only $10$ iterations of this procedure
are necessary for convergence. The resulting median $z$ is taken to be what our observer would measure as the local
Galactic midplane at each $\phi$.
To account for the effect of particle noise, we bootstrap-resample stars within the cylinder $1000$ times and determine
the $95^{\text{th}}$ minus $5^{\text{th}}$ percentile range. This bootstrap resampling is performed by first selecting all stars
within a height of $2\,\kpc$, resampling that selection, and then repeating
the $10$ iterations.

To allow for potential small inaccuracies in the determination of the original fiducial coordinate system,
we also subtract the best fit $\text{A} \sin{\left(\phi + \text{B}\right)} +
\text{C}$ curve from the midplane as a function of azimuth. For $\text{A}$ the
values are $-174$, $62$, and $3.8\,\pc$, for $\text{B}$ the values are $0.3$,
$0.65$, and $0.021\,\text{rad}$ and for $\text{C}$ the values are $-47$, $18$,
and $-15\,\pc$ for m12i, m12f, and m12m, respectively. For the assumed solar
radius of $8.2\,\kpc$, we can approximate the angle offset for the $z$-axis
from the values of $\text{A}$ --- we compute $1.22$, $0.43$, and $0.03\,\deg$
for m12i, m12f, and m12m. These angle offsets are consistent with the values
given in \citet{2018arXiv180610564S} for the difference between the $z$-axis
as defined by the gas and stars.

Figure~\ref{fig:midplane} shows the inferred midplane our imaginary observer
would make as a function of azimuth for each galaxy. The 5--95 percentile range from the bootstrap procedure is shown as the dashed-line error bars. The $90\%$ interquartile
range for each galaxy is $185$, $162$, $84\,\pc$ for m12i, m12f, and m12m.

@@

\subsection{LSR Variations} \label{ssec:lsr_var}
For similar reasons to the midplane varying across the disk, we also expect
that the local standard of rest (LSR) is a function of azimuth as well. We
perform this calculation in Appendix~\ref{app:lsr}, but performing a
best-fit subtraction to correct for misalignment of the original coordinate system (as in the previous section) is more involved, and we
delay to future work. We do note, however, that the variation in the LSR is
less pronounced than for the midplane, at least for how we calculate the LSR.
